---
title: "Influenza Hospitalizations"
author: "[Ryan Beck](https://twitter.com/ryanbeck111)"
output:
  html_document:
    code_folding: hide
---

Last updated: `r format(Sys.time(), '%B %d, %Y')`

This page contains influenza hospital admissions data for several states for use with the [Metaculus FluSight forecasting challenge](https://www.metaculus.com/project/1322/?order_by=close_time&project=1322&status=open).

Note: The data provides previous day hospital admissions, and [consistent with the methodology used for assessing the ground truth data for flusight forecasts](https://github.com/computationalUncertaintyLab/Flusight-forecast-data/tree/master/data-truth) the dates have been shifted one day earlier from those shown in the source data.

Date source: https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/g62h-syeh

Click the "Code" buttons to see the R code used on this page. An R Markdown file of this page is [available here](https://github.com/ryooan/ryooan.github.io/tree/master/forecasting) for anyone who wishes to download and run or modify it themselves.

```{r message=FALSE, warning=FALSE}

library(RSocrata)
library(tidyverse)
library(zoo)
library(scales)
library(kableExtra)
library(reshape2)
library(latticeExtra)

df <- read.socrata("https://healthdata.gov/resource/g62h-syeh.csv")

df <- df %>% mutate(date = as.Date(date,"%Y-%m-%dT00:00:00.000"))

df[is.na(df)] = 0

data_by_date <- df %>% group_by(state, date) %>% summarise(daily = sum(previous_day_admission_influenza_confirmed))

data_by_date <- data_by_date %>% group_by(state) %>% arrange(date) %>% mutate(weekly = rollapply(daily,7,sum, fill = NA, align = "right"))

metaculus_states <- data_by_date %>% filter(state == "NY" | state == "FL" | state == "OK" | state == "CA" | state == "VT" | state == "WY")

#shift the date back 1 since this is previous day reported
metaculus_states <- metaculus_states %>% mutate(date = date - 1)

#GENERATE OVERALL PLOT
state_hospitalizations <- ggplot(metaculus_states, aes(date, weekly)) +
  geom_line(aes(color = state), size=0.75) +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly Influenza Hospitalizations", color="State") +
  #scale_y_continuous(labels = percent, breaks = breaks_width(5)) +
  scale_x_date(breaks="1 month",labels = date_format("%b/%y"), limits = as.Date(c('2020-10-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "lightgrey"))

#GENERATE PLOTS BY STATES
influenza_CA <- metaculus_states %>% filter(state == "CA")
influenza_FL <- metaculus_states %>% filter(state == "FL")
influenza_NY <- metaculus_states %>% filter(state == "NY")
influenza_OK <- metaculus_states %>% filter(state == "OK")
influenza_VT <- metaculus_states %>% filter(state == "VT")
influenza_WY <- metaculus_states %>% filter(state == "WY")

CA_hospitalizations <- ggplot(influenza_CA, aes(date, weekly)) +
  geom_line(size=0.75, color="blue") +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly California Influenza Hospitalizations") +
  scale_y_continuous(breaks = breaks_width(10)) +
  scale_x_date(breaks="1 month",date_minor_breaks="1 week",labels = date_format("%b/%y"), limits = as.Date(c('2021-09-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "gray90"))

FL_hospitalizations <- ggplot(influenza_FL, aes(date, weekly)) +
  geom_line(size=0.75, color="blue") +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly Florida Influenza Hospitalizations") +
  scale_y_continuous(breaks = breaks_width(10)) +
  scale_x_date(breaks="1 month",date_minor_breaks="1 week",labels = date_format("%b/%y"), limits = as.Date(c('2021-09-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "gray90"))

NY_hospitalizations <- ggplot(influenza_NY, aes(date, weekly)) +
  geom_line(size=0.75, color="blue") +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly New York Influenza Hospitalizations") +
  scale_y_continuous(breaks = breaks_width(10)) +
  scale_x_date(breaks="1 month",date_minor_breaks="1 week",labels = date_format("%b/%y"), limits = as.Date(c('2021-09-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "gray90"))

OK_hospitalizations <- ggplot(influenza_OK, aes(date, weekly)) +
  geom_line(size=0.75, color="blue") +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly Oklahoma Influenza Hospitalizations") +
  scale_y_continuous(breaks = breaks_width(10)) +
  scale_x_date(breaks="1 month",date_minor_breaks="1 week",labels = date_format("%b/%y"), limits = as.Date(c('2021-09-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "gray90"))

VT_hospitalizations <- ggplot(influenza_VT, aes(date, weekly)) +
  geom_line(size=0.75, color="blue") +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly Vermont Influenza Hospitalizations") +
  scale_y_continuous(breaks = breaks_width(2)) +
  scale_x_date(breaks="1 month",date_minor_breaks="1 week",labels = date_format("%b/%y"), limits = as.Date(c('2021-09-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "gray90"))

WY_hospitalizations <- ggplot(influenza_WY, aes(date, weekly)) +
  geom_line(size=0.75, color="blue") +
  labs(x="Date", y="Weekly Hospitalizations", title="Weekly Wyoming Influenza Hospitalizations") +
  scale_y_continuous(breaks = breaks_width(10)) +
  scale_x_date(breaks="1 month",date_minor_breaks="1 week",labels = date_format("%b/%y"), limits = as.Date(c('2021-09-01',NA))) +
  #expand_limits(x = as.Date("2021-10-1")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(panel.grid.major = element_line(colour = "grey"), panel.grid.minor = element_line(colour = "gray90"))

#GENERATE TABLE
metaculus_table <- pivot_wider(data = metaculus_states,id_cols = date,names_from = state,values_from = c("daily","weekly"))

metaculus_table <- metaculus_table %>% arrange(desc(date))

state_table_output <- kbl(metaculus_table,col.names = c("Date","CA","OK","WY","FL","NY","VT","CA","OK","WY","FL","NY","VT")) %>%
  add_header_above(c(" " = 1,"Daily Hospitalizations" = 6,"Weekly Hospitalizations" = 6)) %>%
  kable_minimal(full_width = F) %>%
  column_spec(1,width_min = "6em") %>%
  scroll_box(width = "100%", height = "500px")

```


# Overall Time Series of Weekly Hospital Admissions
```{r message=FALSE, warning=FALSE}
state_hospitalizations

```

# Weekly Hospital Admissions for Each State
```{r message=FALSE, warning=FALSE}
CA_hospitalizations
OK_hospitalizations
WY_hospitalizations
FL_hospitalizations
NY_hospitalizations
VT_hospitalizations

```

# Data Table
```{r message=FALSE, warning=FALSE}
state_table_output

```
