---
title: "New York Evictions"
author: "Ryan Beck"
output:
  html_document:
    code_folding: hide
---
Last updated `r format(Sys.time(), '%B %d, %Y')`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The plots and tables below show weekly and cumulative evictions in New York. The data comes from Eviction Lab. The series "2012-2016 Average" represents the evictions for a given week averaged over the 2012 to 2016 period.

Date source: https://evictionlab.org/eviction-tracking/get-the-data/

Information has been independently calculated for use in forecasting New York evictions on Good Judgement Open: https://www.gjopen.com/questions/2063-how-many-new-york-city-eviction-filings-will-be-reported-for-2021

Click the code buttons to view the R code used.
```{r}
# Sample R code to plot weekly filings
# For Milwaukee until week 24 (06/13/2020)

library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(lubridate)
library(kableExtra)

newyork_weekly_2020_2021 <- read.csv("https://evictionlab.org/uploads/newyork_weekly_2020_2021.csv")

NY_data <- newyork_weekly_2020_2021 %>%   
  group_by(week, week_date) %>% 
  summarize(filings_2020 = sum(filings_2020),
            filings_avg = sum(filings_avg, na.rm = T)) %>%
  pivot_longer(cols = filings_2020:filings_avg,
               names_to = "year",
               values_to = "filings",
               names_prefix = "filings_") %>% 
  mutate(year = recode(year,
                       avg = "2012-2016"))

NY_data <- NY_data %>% mutate(week_date=as.Date(week_date, format = "%Y-%m-%d"))

weekly_plot <-  NY_data %>% ggplot(aes(x = week_date,
             y = filings)) +
  geom_line(aes(color = year, group = year)) +
  labs(x = "Date", y = "Weekly Filings", title = "New York Weekly Eviction Filings") +
  scale_x_date(breaks="1 month",labels = date_format("%b/%y")) +
  scale_color_discrete("Period", labels = c("2012-2016 Average","2020-2021 Weekly")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

NY_weekly <- newyork_weekly_2020_2021 %>% group_by(Week = week_date) %>% summarise(Filings = sum(filings_2020),Average = sum(filings_avg))

NY_Intermediate <- NY_weekly %>% separate(Week,c("Year","Month","Day"),remove = F)

Week_Num <- data.frame()
Year_Mod <- data.frame()

for (i in 1:nrow(NY_Intermediate)) {
    Week_Num <- rbind(Week_Num,round(52*(i/52-floor((i-1)/52))))
    Year_Mod <- rbind(Year_Mod,2020+floor((i-1)/52))
}
colnames(Week_Num) <- "Week_Number"
colnames(Year_Mod) <- "Year_Mod"
NY_Overall <- cbind(Week_Num,Year_Mod,NY_Intermediate)

NY_Overall <- NY_Overall %>% group_by(Year_Mod) %>% mutate(Cumulative=cumsum(Filings))

month <- seq(as.Date("2020-01-01"), 
             as.Date("2020-12-01"), 
             by = "1 month")

month_numeric <- lubridate::yday(month) / 365 * 52 + 1
month_label <- lubridate::month(month, label = TRUE)

cumulative_plot <-  NY_Overall %>% ggplot(aes(x = Week_Number,
             y = Cumulative)) +
  geom_line(aes(color = as.factor(Year_Mod), group = as.factor(Year_Mod))) +
  labs(x = "Date", y = "Cumulative Filings", title = "New York Cumulative Eviction Filings") +
  scale_x_continuous(breaks = month_numeric, labels = month_label) +
  scale_color_discrete("Year", labels = c("2020","2021"))

NY_Table <- NY_Overall %>% ungroup %>% select(-Year_Mod,-Year,-Month,-Day)

NY_Table <- NY_Table %>% mutate(across(is.numeric, ~ round(., 0)))

col_order <- c("Week_Number", "Week", "Average",
               "Filings", "Cumulative")
NY_Table <- NY_Table[, col_order]

weekly_plot
```


```{r}
cumulative_plot
```

```{r}
kbl(NY_Table,col.names = c("Week #","Week","2012-2016 Average","Filings","Year to Date")) %>%
  kable_minimal(full_width = F) %>%
  scroll_box(width = "75%", height = "500px")
```